What We Eval in the Shadows

### 2021-04-14 paper/main.tex
m
Fix typos up to Section 4

%% Aviral Goel
%% Pierre Donat-Bouillud
%% Filip Krikava
%% Jan Vitek
%% What is the right order?

\author{Author }
\authornote{Both authors contributed equally to this research.}
\email{a@b.com}
\orcid{}
\author{Author}
\authornotemark[1]
\email{r@s.com}
\affiliation{\institution{Institute} \country{Country}}

@@

\citet{ecoop12} gave a programming language-centric overview of the R language.
They characterized it as a lazy, vectorized, functional language with a rich
complement of dynamic features expressive enough to layer several object systems
on top of the core language. Most data types are sequences of primitive values
constructed by calling the combination function \c{c}, for instance
\c{c("Ha","bye")} evaluates to a vector of two strings, constants such as \c{42}
are vectors of length one. To enable equational reasoning, values accessible
through multiple aliases are copied when written to. Furthermore, values can be
tagged by attributes, these are key-value pairs. For instance, the attribute
\c{dim}-\c{c(2,2)} can be attached to the value at \c{x} by the call
\c{attr(x,"dim")$\leftarrow$c(2,2)}. In this case, the addition of this
attribute turns \c x into a matrix. The \c{class} attribute gives a 'class', in
the object-oriented sense, to a value. So, \c{class(x)<-"human"} sets the class
of \c{x} to \c{human}; classes are used for method dispatch. Every linguistic
construct is desugared to a function call, even control flow statements,
assignments, and bracketing. All functions can be shadowed or redefined, making
the language at the same time remarkably flexible and exceeding challenging to
compile statically as vividly detailed by~\citet{dls19}. R uses a relaxed
call-by-need convention for passing arguments to functions. Each argument is a
thunk composed of an expression, its environment, and a slot for the result,
these are called \emph{promises}. To get the value of an argument, the
corresponding promise must be forced, once forced the promise's result is cached
for future use.

@@

\begin{figure}[!h]\center\small\begin{tabular}{r@{~}l|l|l|l}\hline
\tiny\sc Language&&\sc\tiny Input&\sc\tiny Scope&\tiny\sc Reflective operations\\\hline
\bf Julia&\cite{julia}     & expression& toplevel         & data\\
\bf Java&\cite{cl}  & bytecode  & classloader       & data\\
\bf JavaScript&\cite{ecoop11}& text      & current, toplevel& data\\
\bf R&\cite{R96}  & expression& programmatic      & data, stack, environment\\\hline
\end{tabular}\caption{Design space}\label{comp}
\end{figure}

@@

Given the above, the claim that R is amongst the languages with the most
powerful \eval seems plausible. The rationale for R's design seems to have been
to expose as much of the language and its internals as possible in order to
maximize expressivity. In R, \eval is key tool to extend the language and
implement DSLs, it is also a replacement for macros. By contrast, the designers
of Julia chose to limit \eval. In Julia, only global variables can be
side-effected and environments cannot be readily manipulated. This is designed
to shield optimized code from some of the most pernicious uses of the
facility~\cite{oopsla18a}. Furthermore, Julia provides a versioning
mechanism, called world age, to ensure that any methods defined within an \eval
only become visible at well-defined program points and thus that recompilation
does not have to occur when optimized code is running~\cite{oopsla20a}.

@@

\begin{lstlisting}
 eval <- function(e, envir = parent.frame(),
                   encl = if(is.list(envir)) parent.frame() else baseenv()) ...
\end{lstlisting}
The most general is \eval. Parameter \c e is the \c{expression} to evaluate,
\c{envir} is the evaluation environment, and \c{encl} is used to look up
variables not found in \c{envir}. The default value of latter is either the
caller's environment or top level. \Eval can take several types for \c e, for
our purposes on the \c{expression}. These can be thought of as the abstract
syntax trees returned by the parser. An \c{expression} can be obtained by
calling \c{quote} or \c{parse} with a string:
\begin{lstlisting}
 > quote(a + b)
 # a + b
 > parse(text="a+b")
 # expression(a+b)
\end{lstlisting}
Usually, expressions are obtained from promises using \c{substitute}. Consider a
function called with \c{x = a+b}, then \c f returns an expression object.
\begin{lstlisting}
 > f <- function(x) substitute(x)
 > f(a+b)
 # expression(a+b)
 > substitute(a+b,list(a=1,b=2))
 # 1+2
\end{lstlisting}
The call to \c{substitute(x)} extracts the unevaluated expression from the
promise \c x. Function \c{substitute(e,envir)} takes two arguments, the second
is used to substitute free variables in \c e.
\begin{lstlisting}
 > environment()
 # <environment: R_GlobalEnv>
\end{lstlisting}
The \c{environment} function returns the current environment. Environments nest,
each has a parent. When creating a new environment with \c{new.env}, the parent
is the current environment. Environment chains can be traversed with
\c{parent.env}, until \c{emptyenv} is reached. The top level environment is
\c{.GlobalEnv}, it has parents that represent the packages that have been
loaded. One can also directly read, modify or create new bindings, given any
environment:
\begin{itemize}
\item \c{envir\$v} and \c{get("v",envir=envir)}: read  \c{v} from \c{envir};
\item \c{envir\$v<-2} and \c{assign("v",2,envir=envir)}: store 2 in \c{v}, if
  not found, \c{v} is created in \c{envir}.
\end{itemize}
\noindent Environments are used as hash maps as they have reference semantics
and a built-in string look up. Functions \c{parent.frame} and \c{sys.frame}
return environments further up the call stack.
 \begin{lstlisting}
  evalq <- function(e,envir,encl) eval(quote(e),envir,encl)
  eval.parent <- function(e,n) eval(e,parent.frame(n))
  local <- function(e,envir,encl) evalq(e,new.env())
\end{lstlisting}
The three variants can be expressed with \eval: \c{evalq} quotes its
argument to prevent evaluation in the current environment,
\c{eval.parent(e,n)} evaluates \c{e} in the environment of the {\tt n}-th
caller of this function; finally, \c{local} evaluates \c{e} in a new environment
to avoid polluting the current one.

@@

\citet{wang} analyzed use of dynamic features in 18 Python programs to find if
they affect file change-proneness. Files with dynamic features are significantly
more likely to be the subject of changes, than other files. Chen et al. looked
at the correlation between code changes and dynamic features, including \eval,
in 17 Python programs~\cite{chen}. They did not observe many uses of \eval.
\citet{oscar} performed an empirical study of the usage of dynamic features in
1,000 Smalltalk projects. While \eval itself is not present, Smalltalk has a
rich reflective interface. The authors found that reflective are used in less
than 2\% of methods. The most common reflective method is \c{perform:}; it send
a message that is specified by a string. These features are mostly used in the
core libraries.

@@

\mypara{CRAN.} The Comprehensive R Archive Network
(\small{\url{cran.r-project.org}}) is the largest curated repository of R
packages. It hosts \CranAvailablePackagesRnd packages with 6 new ones submitted
daily~\cite{Ligges2017}. Packages are authored by experienced developers and
abide by well-formedness rules automatically checked on each commit.
Each package comes with sample data. There are three sources of
runnable code in a package: \emph{tests, examples} and \emph{vignettes} --
respectively, unit tests, code snippets from the documentation, and long-form
use-cases written in RMarkdown. Examples and vignettes can be turned into
scripts by extracting the relevant code. From \CranPackages packages we
extracted \CranRunnableScripts scripts that contain \CranRunnableCode lines of
code (\CranRunnableCodeExamplesRnd in examples, \CranRunnableCodeVignettesRnd in
vignettes and \CranRunnableCodeTestsRnd in tests).

\mypara{Kaggle.} The Kaggle website (\small{\url{kaggle.com}}) is an online
platform for data science. It allows users to submit problems and compete to
solve them. Solutions, called \emph{kernels}, are uploaded as plain scripts or
notebooks. The quality of code on Kaggle is not uniform. Each kernel is
runnable, and thus there is one script per kernel. Input data is provided. We
obtained \KaggleKernels kernels. Since Kaggle is not curated, used SHA-1 hashes
to identify and remove \KaggleDuplicates duplicate entries. The remaining
\KaggleUnique kernels are unique solutions to \KaggleCompetitions competition.
They contain \KaggleCode lines of R code. Only \KaggleWithEvals kernels call
\eval.

@@

\medskip
\begin{compactenum}
\item \emph{Download.} Packages are downloaded from CRAN. For scripts, a web
  crawler retrieves code and the Kaggle command line tool gets data.
  Installation is complicated by native dependencies which are not properly
  documented and thus hard to automatically resolve.
\item \emph{Extract.} Given installed packages and scripts, the next step is to
  create runnable programs. The \genthat tool helpfully extracts all runnable
  code snippets from a package and turns each of these into a self-standing
  program~\cite{issta18}. Some Kaggle kernels are already scripts, for those
  nothing more needs be done. Others kernels are packages as notebooks, either
  as Rmarkdown or Jupyter, for those we use \c{knitr} to extract runnable code.
  The body of each extracted program is instrumented with calls our dynamic
  analyzer to ensure that we only record calls to \eval from the code of
  interest and not from bootstrapping or execution harness operations.
\item \emph{Trace.} Each program is executed using our dynamic analysis tool
  which is heavily instrumented interpreter that captures calls to \eval and
  many other fine-grained runtime events. Packages are run twice, once to
  capture \eval calls originating from package code, and a second time to
  capture calls coming from the base libraries. To avoid any interference, each
  program is run in its own process.
  As the GNU R bytecode compiler is written in R, we turn it off
  to avoid recording \eval in its code, and fallback on the interpreter.
\item \emph{Analyze.} Finally, analysis output is merged, cleaned and summarized
  in a post-processing phase driven by series of R scripts. The summarized data
  is then analyzed in RMarkdown notebooks to gather insights. All figures and
  numbers appearing in the paper are generated automatically. Figures are
  produced in PDF by \c{ggplot2}, numbers are exported as \LaTeX macros.
\end{compactenum}

\medskip\noindent The code extraction and tracing steps of the pipeline are run
in parallel~\cite{GNUparallel} orchestrated by a Makefile. Servers have
identical environment thanks to docker images with all dependencies installed.

@@

The tracer registers callbacks to all \eval functions as well as a few other
ones that allow us to better identify where do the expressions come from. For
example, we taint the results of calls to \c{parse} (turns string into
expression) and \c{match.call} (reflects on current call). We also capture calls
to R API for dynamic code loading (\eg\xspace \c{library}, \c{require} and
\c{source}). Next, we subscribe to the events related to a variable definition
and assignment allowing us to record side effects that happen in environments
while evaluating code in \eval. One challenge we encountered during the analysis
was that essentially R only provides source references for top-level calls in a
function body or in the block surrounded by braces. That is to say, a method
whose body would \c{eval(x)} would not have debug information that we can use to
identify that particular \eval. The same expression surrounded by braces would
have a line number and source reference. This is unfortunately not easily fixed
in the R implementation. Instead, we extended the dynamic analysis tool to
attach synthetic source code references to all \eval call sites by traversing
AST. The tracer is implemented as an R package in 3.2K C++ and 1.3K of R code.
For performance reason, most of the tracing is done in C++. While in theory the
implementation is rather straightforward, it is not so in practice:
%
\begin{compactitem}[---]

@@

\item In the dynamic analysis, we run all the \emph{runnable} code we obtained
  from CRAN and Kaggle---\ie real code written by a people with highly varying
  expertise in R and programming in general. This exercise a lot of the corner
  cases of the highly underspecified R behavior.

@@

\mypara{Limitations.} The analysis has some limitation. Even with the extension
described above, there are still \PkgUndefinedRnd \eval calls without source
references. This occurs when a reference to the \eval function is passed as
argument to a higher-order functions or when the \eval call originates from
native code. However, these missing source references account for a meager
\PkgUndefinedRatio of all calls; they are unlikely to affect our results. Other
limitations are that we ignore calls to the native \eval function and to the
alternate \c{rlang::tidy\_eval} function which uses native \eval internally.
The \c{rlang} package introduces a new kind of promise called \c{quosure} which
is evaluated by \c{tidy\_eval}. None of these limitation should invalidate our
conclusions.

@@

There are \PkgEvalCallSites \eval sites in \PkgPackages packages. The proportion
of packages calling \eval is thus \PkgPackagesRatio. Over half of these packages
have fewer than 3 sites, and with the exception \MaxEvalCallSitesPackage which
has \MaxEvalCallSitesCount sites, all packages contain fewer than
\MaxEvalCallSitesRest sites (\cf Fig~\ref{fig:pkgs-eval-callsites-hist} which
shows a histogram of sites per package). These sites appear in \PkgFunsWithEval
functions (\CranFunsWithEvalRatio of all functions in CRAN).

@@

For dynamic analysis, we run \CranRunnableScripts programs extracted from
\CranPackages packages. Any run that does not exercise \eval is discarded. This
left \packageNbruns runs from \packageCorpus packages. There were
\packageAllcalls calls in \packageTriggeredpkgs packages originating from
\PkgHitEvalCallSites unique sites. In terms of coverage, the data exercised
\PkgHitEvalCallSitesAvgRatio of sites, a coverage similar to the code coverage
metric for the packages which is \PkgCodeCoverage. The fact that not all sites
are exercised can be can be chalked down to incomplete tests occasional analysis
failures (\PkgFailedProgramsRatio of programs crashed or timed out).
Fig~\ref{fig:traced-eval-callsites} shows the number of sites that were
exercised; coverage is unequal. Figure~\ref{freq} summarizes the frequency of
dynamic calls to \eval with the left column being the number of calls and the
right, the number of packages in that range. There are \packageFewcalls packages
with low \eval frequency, fewer than 100 calls; and a smaller, but still
significant, number of packages, \packageManycalls to be precise, that use \eval
more than 1,000 times. Package \packageMaxcallspack makes \packageMaxcalls calls
and thus accounts for over half of the observations.
Figure~\ref{tab:variantseval} summarizes the use of variants of \eval. For each
of the four variants, the first rows shows the number of sites in the corpus
(\emph{static}), the second row is the number of sites that were encountered
during analysis (\emph{exercised}), the last row is the number of calls
(\emph{invocations}). The overwhelming majority of sites and calls, are to \eval
itself, \c{eval.parent} is rare, and both \c{evalq} and \c{local} are barely
used at all. The difference between sites and exercised sites underscores the
limitations due to code coverage.

@@

Figure~\ref{cn}(a) shows normalized call counts per site; on the left are
average number of calls from a given site and given run, on the right are counts
of sites that fall in that range. For instance, \packageRunbinh sites are
invoked 2,000+ times per run. Larger numbers suggest loops or recursive contexts
-- this seems to be the exception as most \evals, \packageRunbina, are exercised
50 times or less. Figure~\ref{cn}(b) zooms in on low frequency sites. The x-axis
shows normalized calls and the y-axis is the number sites for that value. Most
low-frequency \evals are invoked only once, about half as many are invoked
twice; after that the frequency quickly drops.

@@

To estimate how much executable code is injected through \eval, we measure the
the number of nodes in the expressions; for example, \c{x+1} counts as 3. We
also measured string lengths of unparsed expressions, but these measurements
were dominated by the size of data objects which could range in the MBs. The
median argument size is \packageMedianszeval (due to the symbols) and the
average is \packageAvgszeval nodes. The largest \eval input observed is
\packageMaxszeval\,-- a significant chunk of code.
Figure~\ref{fig:sizedistribution} shows the distribution of sizes for arguments
of fewer or equal to 25 nodes. The x-axis is the size of arguments in number of
nodes, and the y-axis is the count of arguments with that size. The size drops
rapidly, with few observations larger than 15 nodes. The long tail is omitted
for legibility.

To estimate the work performed in \evals, we count instructions executed by
interpreter. Most invocations perform relatively little work with
\packageSmalleventspct of \evals executing 50 or fewer instructions. The violin
plot of Figure~\ref{ev}(a) corresponds to \evals executing $\leq$ 50
instructions, it is dominated by trivial symbol look ups. Figure~\ref{ev}(b) has
the work-intensive \evals which go all the way to \packageMaxeventsRnd
instructions.

@@

The challenge dealing with Base is that every program uses it and it can thus
generate extreme amounts of data but the data is quite predictable. For the
rest of the paper, we only mention base when there are surprising
observations.

@@

\subsection{Kaggle}
Only \kaggleNbruns scripts trigger \eval. In total, \kaggleAllcalls \eval were
recorded; none are calls to variants. Out of \kaggleStaticeval call sites,
\kaggleTriggeredeval were exercised. Most call sites are invoked only once. Only
one site is called \kaggleMaxcalls times. Expressions amount for
\kaggleCodepercent of arguments. Unlike CRAN corpus, there are very few symbols
(\kaggleSymbolpercent). Most expressions result from calls to \c{parse}, thus
most \evals start with strings. The median argument size is \kaggleMedianszeval,
Which makes sense, as few arguments are only symbols. The largest argument is
\kaggleMaxszeval. The distribution of instructions per eval is similar to CRAN;
\kaggleSmalleventspct of \evals execute fewer than 50 instructions. Manual
inspection of \eval usage in Kaggle suggest that it is consistent with the data
obtained for CRAN. We do not discuss it further.

>>>>

\author{Author }
\authornote{Both authors contributed equally to this research.}
\email{a@b.com}\orcid{}\author{Author}\authornotemark[1]
\email{r@s.com}
\affiliation{\institution{Institute} \country{Country}}

@@

\citet{ecoop12} gave a programming language-centric overview of the R language.
They characterized it as a lazy, vectorized, functional language with a rich
complement of dynamic features expressive enough to layer several object systems
on top of the core language. Most data types are sequences of primitive values
constructed by calling the combination function \c{c}, for instance
\c{c("Ha","bye")} evaluates to a vector of two strings, constants such as \c{42}
are vectors of length one. To enable equational reasoning, values accessible
through multiple aliases are copied when written to. Furthermore, values can be
tagged by attributes, these are key-value pairs. For instance, the attribute
\c{dim}-\c{c(2,2)} can be attached to the value at \c{x} by the call
\c{attr(x,"dim")$\leftarrow$c(2,2)}. In this case, the addition of this
attribute turns \c x into a matrix. The \c{class} attribute gives a 'class', in
the object-oriented sense, to a value. So, \c{class(x)<-"human"} sets the class
of \c{x} to \c{human}; classes are used for method dispatch. Every linguistic
construct is desugared to a function call, even control flow statements,
assignments, and bracketing. All functions can be shadowed or redefined, making
the language at the same time remarkably flexible and exceedingly challenging to
compile statically as vividly detailed by~\citet{dls19}. R uses a relaxed
call-by-need convention for passing arguments to functions. Each argument is a
thunk composed of an expression, its environment, and a slot for the result,
these are called \emph{promises}. To get the value of an argument, the
corresponding promise must be forced, once forced the promise's result is cached
for future use.

@@

\begin{figure}[!h]\center\small\begin{tabular}{r@{~}l|l|l|l}\hline
\tiny\sc Language&&\sc\tiny Input&\sc\tiny Scope&\tiny\sc Reflective operations\\\hline
\bf Julia&\cite{julia}     & expression& toplevel         & data\\
\bf Java&\cite{cl}  & bytecode  & classloader       & data\\
\bf JavaScript&\cite{ecoop11}& text      & current, toplevel& data\\
\bf R&\cite{R96}  & expression& programmatic      & data, stack, environment\\\hline
\end{tabular}\caption{Design space of \eval}\label{comp}
\end{figure}

@@

Given the above, the claim that R is amongst the languages with the most
powerful \eval seems plausible. The rationale for R's design seems to have been
to expose as much of the language and its internals as possible in order to
maximize expressivity. In R, \eval is a key tool to extend the language and
implement DSLs, it is also a replacement for macros. By contrast, the designers
of Julia chose to limit \eval. In Julia, only global variables can be
side-effected and environments cannot be readily manipulated. This is designed
to shield optimized code from some of the most pernicious uses of the
facility~\cite{oopsla18a}. Furthermore, Julia provides a versioning
mechanism, called world age, to ensure that any methods defined within an \eval
only become visible at well-defined program points and thus that recompilation
does not have to occur when optimized code is running~\cite{oopsla20a}.

@@

\begin{lstlisting}
 eval <- function(e, envir = parent.frame(),
                   encl = if(is.list(envir)) parent.frame() else baseenv()) ...
\end{lstlisting}
The most general is \eval. Parameter \c e is the \c{expression} to evaluate,
\c{envir} is the evaluation environment, and \c{encl} is used to look up
variables not found in \c{envir}. The default value of the latter is either the
caller's environment or the base environment.\footnote{It contains all the standard operators and functions in R.} \Eval can take several types for \c e, for
our purposes on the \c{expression}. These can be thought of as the abstract
syntax trees returned by the parser. An \c{expression} can be obtained by
calling \c{quote} or \c{parse} with a string:
\begin{lstlisting}
 > quote(a + b)
 # a + b
 > parse(text="a+b")
 # expression(a+b)
\end{lstlisting}
Usually, expressions are obtained from promises using \c{substitute}. Consider a
function called with \c{x = a+b}, then \c f returns an expression object.
\begin{lstlisting}
 > f <- function(x) substitute(x)
 > f(a+b)
 # expression(a+b)
 > substitute(a+b,list(a=1,b=2))
 # 1+2
\end{lstlisting}
The call to \c{substitute(x)} extracts the unevaluated expression from the
promise \c x. Function \c{substitute(e,envir)} takes two arguments, the second
is used to substitute free variables in \c e.
\begin{lstlisting}
 > environment()
 # <environment: R_GlobalEnv>
\end{lstlisting}
The \c{environment} function returns the current environment. Environments nest,
each has a parent. When creating a new environment with \c{new.env}, the parent
is the current environment. Environment chains can be traversed with
\c{parent.env}, until \c{emptyenv} is reached. The top-level environment is
\c{.GlobalEnv}, it has parents that represent the packages that have been
loaded. One can also directly read, modify or create new bindings, given any
environment:
\begin{itemize}
\item \c{envir\$v} and \c{get("v",envir=envir)}: read  \c{v} from \c{envir};
\item \c{envir\$v<-2} and \c{assign("v",2,envir=envir)}: store 2 in \c{v}, if
  not found, \c{v} is created in \c{envir}.
\end{itemize}
\noindent Environments are used as hash maps as they have reference semantics
and a built-in string look up. Functions \c{parent.frame} and \c{sys.frame}
return environments further up the call stack.
 \begin{lstlisting}
  evalq <- function(e,envir,encl) eval(quote(e),envir,encl)
  eval.parent <- function(e,n) eval(e,parent.frame(n))
  local <- function(e,envir,encl) evalq(e,new.env())
\end{lstlisting}
The three variants can be expressed with \eval: \c{evalq} quotes its
argument to prevent evaluation in the current environment,
\c{eval.parent(e,n)} evaluates \c{e} in the environment of the {\tt n}-th
caller of this function; finally, \c{local} evaluates \c{e} in a new environment
to avoid polluting the current one.

@@

\citet{wang} analyzed use of dynamic features in 18 Python programs to find if
they affect file change-proneness. Files with dynamic features are significantly
more likely to be the subject of changes, than other files. Chen et al. looked
at the correlation between code changes and dynamic features, including \eval,
in 17 Python programs~\cite{chen}. They did not observe many uses of \eval.
\citet{oscar} performed an empirical study of the usage of dynamic features in
1,000 Smalltalk projects. While \eval itself is not present, Smalltalk has a
rich reflective interface. The authors found that reflective methods are used in less
than 2\% of methods. The most common reflective method is \c{perform:}; it send
a message that is specified by a string. These features are mostly used in the
core libraries.

@@

\mypara{CRAN.} The Comprehensive R Archive Network
({\small \url{cran.r-project.org}}) is the largest curated repository of R
packages. It hosts \CranAvailablePackagesRnd packages with 6 new ones submitted
daily~\cite{Ligges2017}. Packages are authored by experienced developers and
abide by well-formedness rules automatically checked on each commit.
Each package comes with sample data. There are three sources of
runnable code in a package: \emph{tests, examples} and \emph{vignettes} --
respectively, unit tests, code snippets from the documentation, and long-form
use-cases written in RMarkdown. Examples and vignettes can be turned into
scripts by extracting the relevant code. From \CranPackages packages we
extracted \CranRunnableScripts scripts that contain \CranRunnableCode lines of
code (\CranRunnableCodeExamplesRnd in examples, \CranRunnableCodeVignettesRnd in
vignettes and \CranRunnableCodeTestsRnd in tests).

\mypara{Kaggle.} The Kaggle website ({\small \url{kaggle.com}}) is an online
platform for data science. It allows users to submit problems and compete to
solve them. Solutions, called \emph{kernels}, are uploaded as plain scripts or
notebooks. The quality of code on Kaggle is not uniform. Each kernel is
runnable, and thus there is one script per kernel. Input data is provided. We
obtained \KaggleKernels kernels. Since Kaggle is not curated, we used SHA-1 hashes
to identify and remove \KaggleDuplicates duplicate entries. The remaining
\KaggleUnique kernels are unique solutions to \KaggleCompetitions competitions.
They contain \KaggleCode lines of R code. Only \KaggleWithEvals kernels call
\eval.

@@

\medskip
\begin{compactenum}
\item \emph{Download.} Packages are downloaded from CRAN. For scripts, a web
  crawler retrieves code and the Kaggle command line tool gets data.
  Installation is complicated by native dependencies which are not properly
  documented and thus hard to automatically resolve.
\item \emph{Extract.} Given installed packages and scripts, the next step is to
  create runnable programs. The \genthat tool helpfully extracts all runnable
  code snippets from a package and turns each of these into a self-standing
  program~\cite{issta18}. Some Kaggle kernels are already scripts, for those
  nothing more needs be done. Others kernels are packages as notebooks, either
  as Rmarkdown or Jupyter, for those we use \c{knitr} to extract runnable code.
  The body of each extracted program is instrumented with calls to our dynamic
  analyzer to ensure that we only record calls to \eval from the code of
  interest and not from bootstrapping or execution harness operations.
\item \emph{Trace.} Each program is executing using our dynamic analysis tool
  which is a heavily instrumented interpreter that captures calls to \eval and
  many other fine-grained runtime events. Packages are run twice, once to
  capture \eval calls originating from package code, and a second time to
  capture calls coming from the base libraries. To avoid any interference, each
  program is run in its own process.
  As the GNU R bytecode compiler is written in R, we turn it off
  to avoid recording \eval in its code, and fallback on the interpreter.
\item \emph{Analyze.} Finally, analysis output is merged, cleaned and summarized
  in a post-processing phase driven by series of R scripts. The summarized data
  is then analyzed in RMarkdown notebooks to gather insights. All figures and
  numbers appearing in the paper are generated automatically. Figures are
  produced in PDF by \c{ggplot2}, numbers are exported as \LaTeX macros.
\end{compactenum}

\medskip\noindent The code extraction and tracing steps of the pipeline are run
in parallel~\cite{GNUparallel} orchestrated by a Makefile. Servers have
identical environments thanks to docker images with all dependencies installed.

@@

The tracer registers callbacks to all \eval functions as well as a few other
ones that allow us to better identify where the expressions come from. For
example, we taint the results of calls to \c{parse} (turns string into
expression) and \c{match.call} (reflects on current call). We also capture calls
to R API for dynamic code loading (\eg\xspace \c{library}, \c{require} and
\c{source}). Next, we subscribe to the events related to a variable definition
and assignment allowing us to record side effects that happen in environments
while evaluating code in \eval. One challenge we encountered during the analysis
was that essentially R only provides source references for top-level calls in a
function body or in the block surrounded by braces. That is to say, a method
whose body would \c{eval(x)} would not have debug information that we can use to
identify that particular \eval. The same expression surrounded by braces would
have a line number and source reference. This is unfortunately not easily fixed
in the R implementation. Instead, we extended the dynamic analysis tool to
attach synthetic source code references to all \eval call sites by traversing
ASTs. The tracer is implemented as an R package in 3.2K C++ and 1.3K of R code.
For performance reason, most of the tracing is done in C++. While in theory the
implementation is rather straightforward, it is not so in practice:
%
\begin{compactitem}[---]

@@

\item In the dynamic analysis, we run all the \emph{runnable} code we obtained
  from CRAN and Kaggle---\ie real code written by people with highly varying
  expertise in R and programming in general. This exercises a lot of the corner
  cases of the highly underspecified R behavior.

@@

\mypara{Limitations.} The analysis has some limitation. Even with the extension
described above, there are still \PkgUndefinedRnd \eval calls without source
references. This occurs when a reference to the \eval function is passed as
argument to a higher-order functions or when the \eval call originates from
native code. However, these missing source references account for a meager
\PkgUndefinedRatio of all calls; they are unlikely to affect our results. Other
limitations are that we ignore calls to the native \eval function and to the
alternate \c{rlang::tidy\_eval} function which uses native \eval internally.
The \c{rlang} package introduces a new kind of promise called \c{quosure} which
is evaluated by \c{tidy\_eval}. None of these limitations should invalidate our
conclusions.

@@

There are \PkgEvalCallSites \eval sites in \PkgPackages packages. The proportion
of packages calling \eval is \PkgPackagesRatio. Over half of these packages
have fewer than 3 sites, and with the exception \MaxEvalCallSitesPackage which
has \MaxEvalCallSitesCount sites, all packages contain fewer than
\MaxEvalCallSitesRest sites (\cf Fig~\ref{fig:pkgs-eval-callsites-hist} which
shows a histogram of sites per package). These sites appear in \PkgFunsWithEval
functions (\CranFunsWithEvalRatio of all functions in CRAN).

@@

For dynamic analysis, we run \CranRunnableScripts programs extracted from
\CranPackages packages. Any run that does not exercise \eval is discarded. This
left \packageNbruns runs from \packageCorpus packages. There were
\packageAllcalls calls in \packageTriggeredpkgs packages originating from
\PkgHitEvalCallSites unique sites. In terms of coverage, the data exercised
\PkgHitEvalCallSitesAvgRatio of sites, a coverage similar to the code coverage
metric for the packages which is \PkgCodeCoverage. The fact that not all sites
are exercised can be chalked down to incomplete tests and occasional analysis
failures (\PkgFailedProgramsRatio of programs crashed or timed out).
Fig~\ref{fig:traced-eval-callsites} shows the number of sites that were
exercised; coverage is unequal. Figure~\ref{freq} summarizes the frequency of
dynamic calls to \eval with the left column being the number of calls and the
right, the number of packages in that range. There are \packageFewcalls packages
with low \eval frequency, fewer than 100 calls; and a smaller, but still
significant, number of packages, \packageManycalls to be precise, that use \eval
more than 1,000 times. Package \packageMaxcallspack makes \packageMaxcalls calls
and thus accounts for over half of the observations.
Figure~\ref{tab:variantseval} summarizes the use of variants of \eval. For each
of the four variants, the first rows shows the number of sites in the corpus
(\emph{static}), the second row is the number of sites that were encountered
during analysis (\emph{exercised}), the last row is the number of calls
(\emph{invocations}). The overwhelming majority of sites and calls, are to \eval
itself, \c{eval.parent} is rare, and both \c{evalq} and \c{local} are barely
used at all. The difference between sites and exercised sites underscores the
limitations due to code coverage.

@@

Figure~\ref{cn}(a) shows normalized call counts per site; on the left are
average number of calls from a given site and given run, on the right are counts
of sites that fall in that range. For instance, \packageRunbinh sites are
invoked 2,000+ times per run. Larger numbers suggest loops or recursive contexts
-- this seems to be the exception as most \evals, \packageRunbina, are exercised
50 times or less. Figure~\ref{cn}(b) zooms in on low frequency sites. The x-axis
shows normalized calls and the y-axis is the number of sites for that value. Most
low-frequency \evals are invoked only once, about half as many are invoked
twice; after that the frequency quickly drops.

@@

To estimate how much executable code is injected through \eval, we measure the number of nodes in the expressions; for example, \c{x+1} counts as 3. We
also measured string lengths of unparsed expressions, but these measurements
were dominated by the size of data objects which could range in the MBs. The
median argument size is \packageMedianszeval (due to the symbols) and the
average is \packageAvgszeval nodes. The largest \eval input observed is
\packageMaxszeval\,-- a significant chunk of code.
Figure~\ref{fig:sizedistribution} shows the distribution of sizes for arguments
of fewer or equal to 25 nodes. The x-axis is the size of arguments in number of
nodes, and the y-axis is the count of arguments with that size. The size drops
rapidly, with few observations larger than 15 nodes. The long tail is omitted
for legibility.

To estimate the work performed in \evals, we count instructions executed by the
interpreter. Most invocations perform relatively little work with
\packageSmalleventspct of \evals executing 50 or fewer instructions. The violin
plot of Figure~\ref{ev}(a) corresponds to \evals executing $\leq$ 50
instructions, it is dominated by trivial symbol look ups. Figure~\ref{ev}(b) has
the work-intensive \evals which go all the way to \packageMaxeventsRnd
instructions.

@@

The challenge dealing with Base is that every program uses it. It can thus
generate extreme amounts of data but the data is quite predictable. For the
rest of the paper, we only mention Base when there are surprising
observations.

@@

\subsection{Kaggle}
Only \kaggleNbruns scripts trigger \eval. In total, \kaggleAllcalls \eval were
recorded; none are calls to variants. Out of \kaggleStaticeval call sites,
\kaggleTriggeredeval were exercised. Most call sites are invoked only once. Only
one site is called \kaggleMaxcalls times. Expressions amount for
\kaggleCodepercent of arguments. Unlike CRAN corpus, there are very few symbols
(\kaggleSymbolpercent). Most expressions result from calls to \c{parse}, thus
most \evals start with strings. The median argument size is \kaggleMedianszeval,
which makes sense, as few arguments are only symbols. The largest argument is
\kaggleMaxszeval. The distribution of instructions per eval is similar to CRAN;
\kaggleSmalleventspct of \evals execute fewer than 50 instructions. Manual
inspection of \eval usage in Kaggle suggest that it is consistent with the data
obtained for CRAN. We do not discuss it further.
